---
title: "Spotify Analysis EDA"
output:
  html_document:
    df_print: paged
  markdown::html_format:
    meta:
      css:
      - default
      - slides
      js: slides
date: "2023-10-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning=FALSE,message=FALSE}


library(tidyverse) 
library(ggplot2) 
library(kknn) 
library(corrplot)
library(readr) 
library(rpart) 
library(rpart.plot) 


```


```{r}
library(readr)
spotify <- read.csv("/Users/evabeyebach/Desktop/spotify.csv")
```




### Exploratory Data Anaylsis 

* Different ways that we could look at the Data to answer our questions could be **histograms**, to learn about the distribution of variables, **boxplots**, to find outliers and see which genres are most popular, **scatterplots** to interpret the variables based on `track_popularity` and **correlation matrix** to find out the correlation between variables.

* For the purpose of this analysis we will not use the variables `track_id`, `track_name`, `track_artist`, `track_album_id`, `track_album_name`, `track_abum_release_date`, `playlist_name` nor `playlist_id`. We only want to focus on the sound parameters and genres.

#### Visual histogram exploration to understand the data set.

We will plot some variables to observe its skewiness, distribution, and interpret them.

```{r,warning=FALSE,message=FALSE}
histograms <- names(spotify)[c(4,12:23)]

songs1 <- spotify %>% 
            select(c(histograms)) %>%
            pivot_longer(cols = histograms) 

songs1 %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~name, ncol = 5, scales = 'free') +
  labs(title = 'Audio Feature Pattern Frequency Plots', x = '', y = '') +
  theme_void()


```

**Observations:**

* `Duration` and `Valence` are **normally distributed**.
* `Danceability`, `Enery` and `Loudness` are **left-skewed**.
* `Acousticness`, `Liveness` and `Speechiness` are **right-skewed**.
* `Key` and `mode` are categorical variables.
* In `popularity` the max is around 50-60, and it is **normally distributed**. There are a lot of zero values, possible missing values. Very few songs have a popularity of above 90.


#### Boxplot to plot different genres vs popularity


```{r}
ggplot(spotify,aes(x = playlist_genre, y = track_popularity, fill = playlist_genre)) +geom_boxplot()
```

```{r}
ggplot(spotify,aes(x = playlist_subgenre, y = track_popularity, fill = playlist_subgenre)) +geom_boxplot()
```


**Observations:** 

* We can see that `pop` is the most popular followed by `latin` and `rock`.
* From the subgenre playlist `post-teen pop` is the most popular followed by `dance pop` and `permament wave`.


#### Visualitazion scatterplots based on parameters and popularity

* Now we will plot scatterplots to compare the different parameters to popularity and see which genre has the best popularity based on the different variables. 
* We also want to see on the most popular songs, which parameter best predicts popularity.
* We are only going to plot it for the first 500 most popular songs, because therefore we can differentiate more if a variable affect of a song is poular.

```{r}
# new dataset with some variables
feature_names <- names(spotify)[c(12:13,15,17:19,23)]

# new dataset only displying variables with 500 most popular songs
songs <- spotify %>% 
  arrange(desc(track_popularity)) %>%
  head(n = 500) %>%
  pivot_longer(cols = feature_names) 

#plotting 
songs %>%
  ggplot(aes(x = name, y = value)) +
  geom_jitter(aes(color = playlist_genre)) +
  facet_wrap(~name, ncol = 4, scales = 'free') +
  labs(title = 'Audio Feature Pattern Frequency Plots', x = '', y = '') +
  theme(axis.text.y = element_blank())


```

**Observation**:

* Higher `danceability`, `energy` and `loudness` usually means more popularity.
* Lower `speechiness` and `accousticness` and `instrumentalness` means more popular.
* An average to low `duration` means more popularity.
* After converting `track_popularity` into dummy variable we could also use other models such as **logistic regression**, **SVM** and **Regression Tree**.


#### Visualization Boxplots

We will plot boxplots on every continuous variable to identify outliers.

##### Outliers 
```{r}
ggplot(data = songs1) +
  geom_boxplot(aes(y = value)) + 
  facet_wrap(~name, nrow = 4, scales = "free") +
  coord_flip() +
  ggtitle("Outlier analysis", subtitle = "For different song attributes") +
  theme_void()
```
**Observations:**

From the boxplots we can observe that a lot of variables (`danceability`, `energy`, `loudness`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `duration`) have outliers. Removing them would influence the analysis a lot, so we will create a new dataset and remove the outliers on that dataset.

##### Dataset copy without oultiers
```{r}
spotify_copy <- spotify

```

##### Truncation
Now, we will truncate energy speechiness, acousticness, instrumentalness and liveness.

```{r}
# truncation danceability, energy, speechiness, acousticness, instrumentalness and liveness
spotify_copy$danceability[spotify_copy$danceability <= 0.28] <- 0.28

spotify_copy$energy[spotify_copy$energy <= 0.2] <- 0.2

spotify_copy$speechiness[spotify_copy$speechiness >= 0.27] <- 0.27

spotify_copy$acousticness[spotify_copy$acousticness >= 0.6] <- 0.6

spotify_copy$instrumentalness[spotify_copy$instrumentalness >= 0.015] <- 0.015

spotify_copy$liveness[spotify_copy$liveness >= 0.4] <- 0.4

```

##### Winsorization
Now, we will winsorize loudness, tempo and duration.
```{r}
# winsorization loudness
# Calculate the 5th and 95th percentiles for 'loudness'
lower_bound_loudness <- quantile(spotify_copy$loudness, 0.05, na.rm = TRUE)
upper_bound_loudness <- quantile(spotify_copy$loudness, 0.95, na.rm = TRUE)

# Winsorize the data
spotify_copy$loudness[spotify_copy$loudness < lower_bound_loudness] <- lower_bound_loudness
spotify_copy$loudness[spotify_copy$loudness > upper_bound_loudness] <- upper_bound_loudness
```

```{r}
# winsorization tempo
# Calculate the 5th and 95th percentiles for 'tempo'
lower_bound_tempo <- quantile(spotify_copy$tempo, 0.05, na.rm = TRUE)
upper_bound_tempo <- quantile(spotify_copy$tempo, 0.95, na.rm = TRUE)

# Winsorize the data
spotify_copy$tempo[spotify_copy$tempo < lower_bound_tempo] <- lower_bound_tempo
spotify_copy$tempo[spotify_copy$tempo > upper_bound_tempo] <- upper_bound_tempo


```

```{r}
#winsorize duration
# Calculate the 5th and 95th percentiles for 'duration'
lower_bound_duration_ms <- quantile(spotify_copy$duration_ms, 0.05, na.rm = TRUE)
upper_bound_duration_ms <- quantile(spotify_copy$duration_ms, 0.95, na.rm = TRUE)

# Winsorize the data
spotify_copy$duration_ms[spotify_copy$duration_ms < lower_bound_duration_ms] <- lower_bound_duration_ms
spotify_copy$duration_ms[spotify_copy$duration_ms > upper_bound_duration_ms] <- upper_bound_duration_ms


```

##### Show cleaned dataset

```{r}
knitr::kable(head(spotify[, 1:23]), "simple")
```



**Observations:**

* Histograms helped us look a the distribution of variables, and since a lot pf them where numeric we concluded that a **Linear regression** and **regression tree** might be a good model to predict `track_popularity`.
* Scatterplots helped us see which parameter might influence more our dependent variable. These were: `instrumentalness`, `speechiness` and `loudness`, since the data was more dense.
* boxplots against `popularity` helped us identify most popular genres: `pop`, `latin` and `rock`.
* Popular subgenres were: `post-teen pop`, `dance pop` and `permanent wave`.

* We still need to perform a **correlation matrix** and **modelling** to see which variables best predict `track_popularity` and see if they agree with our plots.

#### Correlation Matrix

First, we will select the variables that we will use for the correlation matrix. We will convert `playlist_genre` to categorical in order to be able to see the correlation.

```{r}

matrix <- spotify %>%  select(track_popularity, danceability, energy, key, loudness, mode, speechiness,acousticness, instrumentalness, liveness, valence, tempo, duration_ms, playlist_genre)

matrix$playlist_genre<-recode(matrix$playlist_genre, 'pop'=0, 'r&b'=1, 'rock'=2, 'latin'=3, 'edm'=4, 'rap'=5)

cor_matrix <- cor(matrix)
print(cor_matrix)

```

```{r}
corrplot(cor(cor_matrix), method="color", type="upper", order="hclust")
```
The overall ranking of the variables that mostly correlate with `track_popularity` are the following:

1. `Duration` (-0.1396)

2. `Instrumentalness` (-0.1244)

3. `Energy` (-0.1036)

4. `Acousticness` (0.0917)

5. `playlist_genre` (-0.0734)

* Comparing to the scatterplots one would think that `instrumentallness`, `speechiness` and `loudness` would have the strongest correlations since they have most points out of the most popular songs in the smallest interval.

* In this correlation matrix however `Duration`, `Instrumentalness` and `Energy` would influence more `track_popularity`.

